---
title: "202hw9"
author: "Schoen"
date: "2023-04-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Reading in Data
```{r}
library(dplyr)
library(tidyverse)
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
```

Part 1- Data Tidying
```{r}
av1 = av %>% filter(Name.Alias != "")

av2 = av1 %>% gather(key = time, value = death, c(11, 13, 15, 17, 19)) %>% 
  select(Name.Alias, time, death) %>%
  mutate(time = parse_number(time))

av3 = av1 %>% gather(key = time, value = return, c(12, 14, 16, 18, 20)) %>% 
  select(Name.Alias, time, return) %>%
  mutate(time = parse_number(time))

avtidy = left_join(av2, av3, by = c("Name.Alias", "time"))
```

Check for multiple rows for the same name
```{r}
avtidy[duplicated(avtidy$Name.Alias),]
```

"Vance Astrovik" has several different rows, so we filter out his information. 
```{r}
av1 = av1 %>% filter(Name.Alias != "Vance Astrovik")

av2 = av1 %>% gather(key = time, value = death, c(11, 13, 15, 17, 19)) %>% 
  select(Name.Alias, time, death) %>%
  mutate(time = parse_number(time))

av3 = av1 %>% gather(key = time, value = return, c(12, 14, 16, 18, 20)) %>% 
  select(Name.Alias, time, return) %>%
  mutate(time = parse_number(time))

avtidy = left_join(av2, av3, by = c("Name.Alias", "time"))
```

Part Two:

Checking Statement that there is a 2/3 chance the Avenger returned from their first stint in the afterlife. 

```{r}
avtidy %>% filter(time == 1, death == "YES") %>% 
  summarise(total = n(),
            return = sum(return == "YES"), 
            noreturn = sum(return == "NO"), 
            missing = sum(return == ""))
```
As we can see 44/63 Avengers returned from their first death. This turns out to be 69.84% of Avengers returned from their first death, which is close to the stated 66.67%. Thus the statement is verified. 

Check Statement that there is a 50% chance they recovered from a second or third death.

```{r}
avtidy %>% filter(time == 2, death == "YES") %>% 
  summarise(total = n(),
            return = sum(return == "YES"), 
            noreturn = sum(return == "NO"), 
            missing = sum(return == ""))

avtidy %>% filter(time == 3, death == "YES") %>% 
  summarise(total = n(),
            return = sum(return == "YES"), 
            noreturn = sum(return == "NO"), 
            missing = sum(return == ""))
```

For the second death, 8/15 or 53.33% of the Avengers returned. For the third death, 1/2 or 50% of the Avengers returned. Therefore the statement that there is a 50% chance they recovered from a second or third death is verified.



